# Introduction to R
()
- Whenever variables are collected in a `c()`, standard operation will be done element-wise.
- Collection will be recycled if the length does not match, but is a multiplation of eachother.

```{r}
x <- c(TRUE, FALSE, 0, 6)
y <- c(FALSE, TRUE, FALSE, TRUE)
!x
x & y
x && y
x | y
x || y
```

Handy: 
- typeof(), 
- str()
```{r}
x = c(1,2,3, NA)
mean(x, na.rm = TRUE)
typeof(x)

x <- list("a" = 2.5, "b" = TRUE, "c" = 1:3)
str(x)
```

## Type of variables
### Numeric
Numeric <- double (=pythonic float), integer, or complex
```{r}
typeof(5)
typeof(0xb) # hexanomal

typeof(5L)
typeof(5L+5L)
typeof(5L/5L)

typeof(5i)
```

#### Operations
##### Numbers
```{r}
x <- 5
y <- 16

x + y
x - y
x * y
y / x
y %/% x # integer division
y %% x # modulus
y ^ x # exponent
```

##### Vectors

###### Making of
```{r}
seq(1, 3, by=0.2) # specify step size
seq(1, 5, length.out=4) # specify length of the vector
```

###### Operations
```{r}
x <- c(2, 8, 3)
y <- c(6, 4, 1)

# Default is element-wise:
x + y
x > y
```

```{r}
x <- c(2, 1, 8, 3)
y <- c(9, 4)
x + y
x - 1 # 1 is recycled to (1,1,1,1)
x + c(1, 2, 3) # (1,2,3) is recycled to (1,2,3,1) but warning issued
```

###### Slicing/dicing
```{r}
x <- 11:17
# access 3rd element
x[3]    
# access 2nd and 4th element
x[c(2, 4)]     
# access all but 1st element
x[-1]          
# cannot mix positive and negative integers
# x[c(2, -4)] NOTE: <- that gives an error


x <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3, ncol = 3, byrow = TRUE)
x
# select rows 1 & 2 and columns 2 & 3
x[c(1,2),c(2,3)]    

# leaving column field blank will select entire columns
x[c(3,2),]    

#  leaving row as well as column field blank will select entire matrix
x[,]    

# select all rows except first
x[-1,]    
```

### Text
Text <- character
```{r}
typeof('example')
```

Concatinate
```{r}
x = paste('Here', "I", "am")
y = paste0("Rock", "You", "Like", "A", "Hurricane")

x
y
```







if...else if...else statements:
```{r}
student.score <- 90

if        (student.score >= 90) {print('A Grade')
} else if (student.score >= 80) {print('B Grade')
} else                  {print('F Grade')
}

# create a vector
a = c(5,7,2,9)
# check if each element in a is even or odd
ifelse(a > 5.5,"PASS","FAIL")
```

##### for loop
```{r}
for (y in 1:20) {
    if (y %% 2)
        next
        print(y)
}
```


##### Next
Next will skip the iteration to which the condition is applicable to.
```{r}
for ( x in 1:20) {

    if ( x %% 2 )

        next

    print(x)

}
```

##### Function
```{r}
pow <- function(x, y=2) {
# function to print x raised to the power y
result <- x^y
return(paste(x,"raised to the power", y, "is", result))
}

pow(3)
```

Return multiple results
```{r}
multi_return <- function() {
my_list <- list("color" = "red", "size" = 20, "shape" = "round")
return(my_list) 
}
# call function multi_return() and assign the result to variable a
a <- multi_return()
a$color
a$size
a$shape
```


Infix operations
```{r}
a <- 5
b <- 6
`+`(a,b)
```

#### Handy functions!
Sample by default will shuffle a list.
If it is given a max to return, it will return that amount.
```{r}
sample(c("H","T"),10, replace = TRUE)


```